R-LIME: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-02-27 paper/paper.tex
Work on paper

\ifnum\mode=0
	\documentclass{article}
	\usepackage{prmlthesis}
\else
	\documentclass[runningheads]{llncs}
\fi

@@

		{%
			\def\AB#1{\ifnum#1=0 A\else B\fi}
			\def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
			\ifnum\mode=0
				\renewcommand{\arraystretch}{0.99}
			\else
				\renewcommand{\arraystretch}{1.02}
			\fi
			\begin{figure}[tbp]
				\foreach\a in {0,1}{%
						\centering
						\begin{subfigure}{\textwidth}
							\ifnum\mode=0
								\small
							\fi
							\centering
							\begin{tabular}{p{14em}m{16em}}
								\toprule
								\csvreader[no head, late after line= \\]{%
									\dir/\index{\a}.csv
								}{}{%
								\ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
								}
								\bottomrule
							\end{tabular}
							\caption{Instance~\AB{\a}}\mylabel{\a}
							\vspace{15pt}
						\end{subfigure}
					}
				\vspace{-15pt}
				\caption[Two instances sampled from recidivism dataset]{%
					Two instances sampled from recidivism dataset.
				}\label{fig:instance}
			\end{figure}
		}
		{%
			\ifnum\mode=0
				\def\scale{0.3}
				\def\imgwidth{0.45\textwidth}
			\else
				\def\scale{0.315}
				\def\imgwidth{0.495\textwidth}
			\fi
			\def\hspacebase{\hspace{-1.5em}}
			\def\vspacebase{\vspace{0.5em}}
			\def\vspacebeforecaption{\vspace{-0.4em}}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Asample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:A-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{1.0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance A by LIME and R-LIME]{%
					Explanation for Instance A by LIME and R-LIME.
				}\label{fig:A}
			\end{figure}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Bsample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:B-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance B by LIME and R-LIME]{%
					Explanation for Instance B by LIME and R-LIME.
				}\label{fig:B}
			\end{figure}
		}
	}
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.

@@

\section{Challenges and Future Perspectives}
 {%
  \def\imgwidth{0.49\textwidth}
  \def\scale{0.08}
  \begin{figure}[t]
	  \centering
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[scale=\scale]{newlime}
		  \caption{R-LIME for balanced label distribution.}\label{fig:balanced}
	  \end{subfigure}
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[scale=\scale]{newlime_for_imbalanced_data}
		  \caption{R-LIME for imbalanced label distribution.
		  }\label{fig:imbalanced}
	  \end{subfigure}
	  \caption[Behavior of R-LIME for balanced and imbalanced label distribution]{%
		  Behavior of R-LIME for balanced and imbalanced label distribution.
		  In case of imbalanced label distribution,
		  the approximation region covers the entire input space and the
		  linear approximation model always outputs the majority label.
	  }
  \end{figure}
 }

@@

\ifnum\mode=0
	\acknowledge
	I would like to express my gratitude to Associate Professor Keigo Kimura
	from the Department of Mathematical Science,
	Division of Information Recognition,
	Graduate School of Information Science and Technology, Hokkaido University.
	He provided invaluable guidance,
	from setting the research theme to outlining the strategy and content of this study.
	Additionally,
	I appreciate the diverse insights shared by Professor Mineichi Kudo in the same laboratory.
	Finally, I am deeply thankful to all the members of the laboratory
	for their helpful advice and collaboration throughout the research process.
\fi

>>>>

\ifnum\mode=0
	\documentclass[11pt]{article}
	\usepackage{prmlthesis}
	\usepackage[a4paper]{geometry}
\else
	\documentclass[runningheads]{llncs}
\fi

@@

		{%
			\def\AB#1{\ifnum#1=0 A\else B\fi}
			\def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
			\ifnum\mode=0
				\renewcommand{\arraystretch}{0.97}
			\else
				\renewcommand{\arraystretch}{1.02}
			\fi
			\begin{figure}[tbp]
				\foreach\a in {0,1}{%
						\centering
						\begin{subfigure}{\textwidth}
							\ifnum\mode=0
								\small
							\fi
							\centering
							\begin{tabular}{p{14em}m{16em}}
								\toprule
								\csvreader[no head, late after line= \\]{%
									\dir/\index{\a}.csv
								}{}{%
								\ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
								}
								\bottomrule
							\end{tabular}
							\caption{Instance~\AB{\a}}\mylabel{\a}
							\vspace{15pt}
						\end{subfigure}
					}
				\vspace{-15pt}
				\caption[Two instances sampled from recidivism dataset]{%
					Two instances sampled from recidivism dataset.
				}\label{fig:instance}
			\end{figure}
		}
		{%
			\ifnum\mode=0
				\def\scale{0.3}
				\def\imgwidth{0.45\textwidth}
				\def\hspacebase{\hspace{-0.5em}}
			\else
				\def\scale{0.315}
				\def\imgwidth{0.495\textwidth}
				\def\hspacebase{\hspace{-1.5em}}
			\fi
			\def\vspacebase{\vspace{0.5em}}
			\def\vspacebeforecaption{\vspace{-0.4em}}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Asample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:A-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{1.0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance A by LIME and R-LIME]{%
					Explanation for Instance A by LIME and R-LIME.
				}\label{fig:A}
			\end{figure}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/lime-\Bsample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:B-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance B by LIME and R-LIME]{%
					Explanation for Instance B by LIME and R-LIME.
				}\label{fig:B}
			\end{figure}
		}
	}
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.

@@

\section{Challenges and Future Perspectives}
 {%
  \def\imgwidth{0.45\textwidth}
  \begin{figure}[t]
	  \centering
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.7\textwidth]{newlime}
		  \caption{R-LIME for balanced label distribution.}\label{fig:balanced}
	  \end{subfigure}
	  \begin{subfigure}[t]{\imgwidth}
		  \centering
		  \includegraphics[width=0.7\textwidth]{newlime_for_imbalanced_data}
		  \caption{R-LIME for imbalanced label distribution.
		  }\label{fig:imbalanced}
	  \end{subfigure}
	  \caption[Behavior of R-LIME for balanced and imbalanced label distribution]{%
		  Behavior of R-LIME for balanced and imbalanced label distribution.
		  In case of imbalanced label distribution,
		  the approximation region covers the entire input space and the
		  linear approximation model always outputs the majority label.
	  }
  \end{figure}
 }

@@

\ifnum\mode=0
	\acknowledge
	I would like to thank Associate Professor Keigo Kimura
	of the Laboratory for Pattern Recognition and Machine Learning,
	Division of Computer Science and Information Technology,
	Graduate School of Information Sci.\ and Tech., Hokkaido University.
	He provided valuable guidance
	from setting the research theme to outlining the research method.
	Furthermore,
	I appreciate the various insights shared by Professor Mineichi Kudo
	in the same laboratory.
	Finally, I am deeply thankful to all the members of our laboratory
	for their helpful advice and cooperation throughout the research process.
\fi
